---
title: "SECOND ASSIGNMENT OF SML"
author: "Eliane Mopo"
date: '2024-01-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(kernlab)
library(ROCR)
library(MASS)
library(lattice)
library(caret)
library(corrplot)
library(rpart)
library(rpart.plot)
library(class)
library(audio)
library(xtable)
```




# 1. Comment on the peculiarities of the dataset from a point of view dimensionality.
```{r}
getwd()
XY <- read.csv('accent-raw-data-1.csv')
X <- as.matrix(XY[,-1])
Y <- XY[,1]

p <- ncol(XY)-1 # Dimensionality of the input space
p
n <- nrow(XY) # Sample size
n
```
# 2. Let us the ts.plot() function to plot speakers {9, 45, 81, 99, 126, 234}.

```{r}
par(mfrow = c(2,3))
speakers <- c(9,45,81,99,126,234)
for(i in speakers)
{
  ts.plot(as.numeric(X[i,]), main=paste('speaker',i) , ylab= 'soundtracks' , xlab='Time')
}
par(mfrow = c(1,1))
```

# 3. Comment comparatively on the features of the plotted soundtracks.
```{r}
for(i in speakers)
{
  play(as.numeric(X[i,]))
  Sys.sleep(3)
}
```

# 2. 

```{r}
accent_mfcc <- read.csv('accent-mfcc-data-1.csv')


accent_mfcc$y <- as.factor(ifelse(accent_mfcc$language=="US",1,0))
xy <- accent_mfcc[,-1]
p <- ncol(xy) # Dimensionality of the input space
x <- xy[,-p]
y<- xy[,p]

```

# 1. Generate separate confusion matrices for each of the three methods
```{r}
n <- nrow(x)
vk <- seq(1, floor(sqrt(n))) # Grid of values of k
nk <- length(vk)             # Number of values of k considered
cv.error <- numeric(nk)     # Vector of cross validation errors for each k
nc <- n                     # Number of observations used for cross validation
c <- 10                   # Number of folds. We are doing c-fold cross validation
S <- sample(sample(nc))     # We randomly shuffle the data before starting CV
m <- ceiling(nc/c)          # Maximum Number of observations in each fold
held.out.set <- matrix(0, nrow=c, ncol=m) # Table used to track the evolution
for(ic in 1:(c-1))
{
  held.out.set[ic,] <- S[((ic-1)*m + 1):(ic*m)]
}

held.out.set[c, 1:(nc-(c-1)*m)] <- S[((c-1)*m + 1):nc]  # Handling last chunk just in case n!=mc
for(j in 1:nk)
{
  for(i in 1:c)
  {  
    out <-  held.out.set[i,]
    yhatc<- knn(x[-out,], x[out,],y[-out],  k=vk[j])
    cv.error[j]<-cv.error[j] + (length(out)-sum(diag(table(y[out],yhatc))))/length(out)
  }
  cv.error[j]<-cv.error[j]/c
}
k <-which.min(cv.error)

```

```{r}
#1NN
kNN.mod <- knn(x, x, y, k=1, prob = TRUE)
conf.mat.kNN <- table(y,kNN.mod)
conf.mat.kNN
```


```{r}
#kNN
kNN.mod <- knn(x, x, y, k=k, prob = TRUE)
conf.mat.kNN <- table(y,kNN.mod)
conf.mat.kNN
```

```{r}
# Trees
tree.mod <- rpart(y ~ ., data =xy, method = "class")
y.pred <- predict(tree.mod, x, type = "class")
conf.mat.tree <- table(y,y.pred )
conf.mat.tree

```


```{r}
# SVM
y.svm <- ksvm(y ~., data = xy, kernel = 'rbfdot', type='C-svc', prob.model=TRUE)
y.pred <- predict(y.svm, x)
conf.mat.svm <- table(y,y.pred )
conf.mat.svm
```

# 2. Plot comparative ROC curves for all the three methods

```{r}
# kNN

prob <- attr(kNN.mod, 'prob')
prob <- 2*ifelse(kNN.mod == "0", 1-prob, prob) - 1
pred.kNN <- prediction(prob, y)
perf.kNN <- performance(pred.kNN, measure='tpr', x.measure='fpr')
#Trees
prob <- predict(tree.mod, x, type='prob')[,2]
pred.tree <- prediction(prob, y)
perf.tree <- performance(pred.tree, measure='tpr', x.measure='fpr')
# SVM

prob     <- predict(y.svm, x, type='probabilities')[,2]
pred.svm <- prediction(prob, y)
perf.svm <- performance(pred.svm, measure='tpr', x.measure='fpr')

# plots of the ROC Curves
plot(perf.kNN, col=1, lwd= 2, lty=1, main=paste('Comparative ROC curves'))
plot(perf.kNN, col=2, lwd= 2, lty=1, add=TRUE)
plot(perf.tree, col=3, lwd= 2, lty=1, add=TRUE)
plot(perf.svm, col=4, lwd= 2, lty=1, add=TRUE)
abline(a=0,b=1)
legend('bottomright', inset=0.05, c('kNN', 'Tree', 'SVM'),col=4:10, lty=1)

#The ROC curve of 1NN will be perfect if you plot it, but here, I decided to plot the ROC of 3NN
```
# 3. Use a 60% − 40% Training-Test set split to generate comparative boxplots of the
#test error based on 100 replications.
```{r}
epsilon <- 0.6            # Proportion of Training set
R <-100                    # Numbers of replications
te.err <- matrix(0,nrow = R, ncol=3)

```

```{r}
for(r in 1:R)
{
  # Split the data
  
  hold <- stratified.holdout(y, epsilon)
  id.tr <- hold$idx1
  id.te <- hold$idx2
  ntr   <- length(id.tr)
  nte   <- length(id.te)
  
  
  # kNN
  yte.hat.kNN <- knn(x[id.tr,], x[id.te,],y[id.tr], k=k, prob=TRUE)
  ind.err.te <- ifelse(y[id.te]!=yte.hat.kNN,1,0)
  te.err[r,1] <- mean(ind.err.te)
  
  # Tree
  tree.mod <- rpart(y ~ ., data = xy[id.tr,], method = "class")
  yte.hat.tree <- predict(tree.mod, x[id.te, ], type='class')
  ind.err.te<- ifelse(y[id.te]!=yte.hat.tree,1,0)  # Random variable tracking error. Indicator
  te.err[r,2]  <- mean(ind.err.te)
  
  # SVM
  svm.mod <- ksvm(y~., data=xy[id.tr, ], kernel='rbfdot', type='C-svc', prob.model=TRUE)
  y.te.hat <- predict(svm.mod, x[id.te, ], type='response')
  ind.err.te     <- ifelse(y[id.te]!=y.te.hat ,1,0)
  te.err[r,3]  <- mean(ind.err.te)
}
```

```{r}
test.error <- data.frame(te.err)
colnames(test.error) <- c('KNN', 'Tree', 'SVM')
boxplot(test.error , col=2:4 , main= 'Comparative boxplots of the
test error based on 100 replications')
```


```{r}
#  4. Comment on the predictive performances.
xtable(summary(test.error))
var(test.error$kNN)
var(test.error$Tree)
var(test.error$SVM)
```

# 5. Reconsider the confusion matrix of the best method and comment on the similarity
#between speaking accents.
